Multimodal Language Learning in the Face of Missing Modalities

### 2022-10-06 Paper/Figures/unused/results-table-old.tex
Paper submitted to TMLR

\documentclass[10pt]{article} % For LaTeX2e
\usepackage{tmlr}
% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{tmlr}
% To de-anonymize and remove mentions to TMLR (for example for posting to preprint servers), instead use the following:
%\usepackage[preprint]{tmlr}

% Optional math commands from https://github.com/goodfeli/dlbook_notation.
\input{math_commands.tex}

\usepackage{hyperref}
\usepackage{url}

% My Packages
\usepackage{cleveref}
\usepackage{comment}
\usepackage{todonotes}
\newcommand{\todokdinline}[1]{\todo[color=red!20,inline]{{KD: \small #1}}}
\newcommand{\todokd}[1]{\todo[color=red!20]{{\small #1 -- Kasra}}}
% \newcommand{\todocainline}[1]{\todo[color=yellow!20,inline]{{CA: \small #1}}}
\newcommand{\todocm}[1]{\todo[color=green!40]{\small #1 -- Cynthia}}
\newcommand{\todocmi}[1]{\todo[inline,color=green!40]{\small #1 -- Cynthia}}
\newcommand{\todoff}[1]{\todo[color=blue!20]{\small #1 -- Frank}}
\newcommand{\todoffinline}[1]{\todo[inline,color=blue!20]{\small #1 -- Frank}}
\newcommand{\todoed}[1]{\todo[color=cyan!20]{\small #1 -- Ed}}

\newcommand{\CitationNeeded}[1]{{\textbf {\color{red}Cite #1}}}
\newcommand{\ProofRead}[1]{{\textbf {\color{red}Proofread please #1}}}
\newcommand{\Complete}[0]{{\textbf {\color{red}Complete it }}}
\newcommand{\Rephrase}[1]{{\textbf {\color{red}Rephrase please #1}}}
\newcommand{\TD}[1]{{\color{red}{\textbf{TODO}}} #1}
\newcommand{\OK}[1]{{\color{green}{\textbf{DONE}}} #1}
\newcommand{\ours}{\textsc{EMMA}}
% \newcommand{\ours}{\textsc{EMMA}}
\newcommand{\geom}{\textsc{Geometric Alignment}}
\newcommand{\supcon}{\textsc{SupCon}}

%%%%%%%%%%%%%%%%%%%%
% EDITING LINK: 
% https://www.overleaf.com/3575751412sjsdmszkxsyh
%%%%%%%%%%%%%%%%%%%%

\title{Multimodal Contrastive Learning for Object Retrieval in the Face of Modality Failures \\ Extended Multimodal Alignment \\ Combining Geometric and classification ... }

% Authors must not appear in the submitted version. They should be hidden
% as long as the tmlr package is used without the [accepted] or [preprint] options.
% Non-anonymous submissions will be rejected without review.

\author{\name Kyunghyun Cho \email kyunghyun.cho@nyu.edu \\
      \addr Department of Computer Science\\
      University of New York
      \AND
      \name Raia Hadsell \email raia@google.com \\
      \addr DeepMind
      \AND
      \name Hugo Larochelle \email hugolarochelle@google.com\\
      \addr Mila, Universit\'e de Montr\'eal \\
      Google Research\\
      CIFAR Fellow}

% The \author macro works with any number of authors. Use \AND 
% to separate the names and addresses of multiple authors.

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\def\month{MM}  % Insert correct month for camera-ready version
\def\year{YYYY} % Insert correct year for camera-ready version
\def\openreview{\url{https://openreview.net/forum?id=XXXX}} % Insert correct link to OpenReview for camera-ready version

\begin{document}

\maketitle

\begin{abstract}
    % Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval. Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment. However, physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks. In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities. We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command. We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive models when modalities are ablated.

    % shorter version

    We propose a generalized geometric combined with cross-entropy loss function that can be used to learn retrieval models that incorporate an arbitrary number of views of a particular piece of data, and compounded by the challenge of retrieval when a modality becomes unavailable. 
    \todoff{this red text can (should) be shortened}
    \todokd{I shortened it. Please remove both notes if it's resolved}
    {\color{red}Our study is motivated by needs in robotics and computer-human interaction, where an agent has many sensors and thus modalities in which a human may interact both to communicate a desired goal and for the agent to recognize a desired target object.
    Tying these various modalities together relates to the task of grounded language understanding. However, there has been little research on integrating more than two modalities.
    % \todokdinline{Duplicate, it's already mentioned in intro}in which natural language is used as a query against objects in a physical environment, allowing a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval.
    % However, current work has neglected the real-world constraint that sensors fail. For practical applications,
    %  we need mechanisms for accurate object retrieval given that a modality may fail without notice.
    }
    % \todoff{There's very little in here about actual method or conclusions}
    There has been widely popular works on self-supervised contrastive learning based on cross-entropy, but there is an entire other approach based on explicit geometric alignment. However, there has been no work on combining the two approaches for multimodal learning to the best of our knowledge. We propose to combine the two approaches and argue that the two are complementary.
    We demonstrate the usability of our model on a grounded language object retrieval scenario, where an intelligent agent has to select an object given an unconstrained language command. 
    We leverage four modalities including vision, depth sensing, text, and speech, and we show that our model can outperform state-of-the-art contrastive learning models by $2\%$ while converging/learning/training $5$ times faster (takes $80\%$ less time or is $20\%$ of the baseline).
    %when modalities are ablated. 
    The code will be publicly available on GitHub and will be included for the camera-ready version (it is redacted for anonymity). 
\end{abstract}

\section{Introduction}
\label{intro}

\todoffinline{nix this first sentence and revise the start of the second. We're not doing any cogsci-y stuff, or even human studies tied to modality deprivation. You could take most of the red text in the abstract I mentioned should be compressed and rework it here}
\todokdinline{I took a shot at nixing it. Please remove both notes if it's resolved. I don't get the vibe of cogsci-y stuff or human studies from the first paragraph, would you please refer to those sentences?}
Humans successfully use multiple sensory inputs to understand, interact with, and retrieve information from the world around them. Inspired by the multimodal nature of human interactions with the world, it is intuitive that agents learning about the world, upon encountering new concepts and new objects, should form a model that incorporates information from all available sensors and data sources. Benefits of integrating multiple modalities is twofold; first, additional modalities can help in the cases when other modalities fail, and second, complementary information can be extracted from different modalities that can help with understanding the world.
% 
Grounded language understanding, in which natural language is used as a query against objects in a physical environment, allows a real-world, intuitive mechanism by which users can instruct physical agents to engage in tasks such as object retrieval. Visuolinguistic approaches to such object inference tasks typically involve training on large pools of image/text pairs and then using language to subselect elements of the sensed environment~\cite{hong2021gilbert,Zhuge_2021_CVPR_VLP}.
\todoffinline{Unless Hong et al is a review paper, add at least one more citation}
\todokdinline{I believe this sentence was added to cite some relevant work from SIGIR, so we can remove it. However, I added one more citation anyway.}
Although physical agents such as robots typically have access to sensory and interactive modalities beyond vision, and learning from multiple modalities can improve performance on downstream tasks, most approaches use at most two sensory inputs (e.g., visual data such as RGB plus depth images) with single labels, such as those provided by textual natural language. Simultaneously using additional inputs from different modalities is an underexplored area, in part due to the domain-specific nature of such \textit{n}-ary learning approaches. With the modern proliferation of audio and text based communication and home agents (e.g., Alexa/Google Home), there is a growing need to handle more modalities, and simultaneously their potential failures.

One difficulty with working with complex multimodal data is the increased likelihood that one or more modalities may have missing information; hardware can become damaged or shipped with defects, sensors can get blocked or obstructed, and various adverse but not uncommon conditions that remove a modality from use.
Current multimodal approaches are typically not robust to the loss of one or more modalities at test time, as may happen if, for example, a physical agent fails to retrieve data from a particular sensor. In order to fully leverage multimodal training data while being robust to missing information, we propose a generalized distance-based loss function that can be extended to learn retrieval models that incorporate an arbitrary number of modalities.

We consider the domain of grounded language-based object retrieval~\cite{hu2016natural, triplet_loss_2021_CVPR}, in which objects in an environment must be identified based on linguistic instructions. This can be considered a special case of image retrieval~\cite{huang2017deep, ma2020large, novak2015large, vo2019composing} in which objects are identified using visual inputs in combination with other sensor modalities. Approaches to acquiring grounded language have explored various combinations of sensor inputs such as depth and RGB with labels provided by textual language or  speech~\cite{RichardsDarvishMatuszekCategoryFree20}. 
However, despite the multisensory nature of object retrieval, much of the existing work has not previously been extended to include an arbitrary number of modalities.

\todoff{revise this goal. It sets up the reviewers to take the easy way out and say results are too incremental}
The high-level goal of this work is to take arbitrary input modalities about novel objects, including both spoken and written language, and build a model that allows a system to correctly retrieve that object given a subset of those input modalities. This is a generalization of approaches to grounded language learning in which specific input modalities are labeled with language to allow for future identification, but explicitly seeks to be agnostic about the nature of the individual sensory and linguistic inputs. Our contributions are as follows:
\begin{enumerate}
    \item Proposing a new approach to multimodal learning by combining geometric and cross-entropy based methods and showing that it outperforms and converges faster compared to each method separately.
    % \item Outperforming state-of-the-art models for grounded language learning by differentiating visual input modalities such as RGB and depth, rather than combining them into a single modality.

    \item Our method can be easily extended to an arbitrary number of modalities.\todoffinline{potential counter: isn't the method quadratic?}
    \todokdinline{Yes, but nonetheless it can be extended.}

    \item Outperforming state-of-the-art contrastive learning~\cite{chen2020simple} and supervised contrastive learning~\cite{NEURIPS2020_supervised_contrastive} models by $2\%$ when all modalities are available, and achieving the peak performance $5$ times faster than both methods. 
    % Ours takes 8 epochs, baselines take 40 epochs.
    \todokdinline{Should we break this contribution into two contributions, one for faster learning and the other for performance compared to both supcon and vanilla contrastive learning?}

    \item Demonstrating robustness in the face of missing information, when one or more modalities are ablated at test time.

    \item Demonstrating the separate utility of speech and text as sources of label information by treating them as sensory input modalities, rather than explicitly as labels.

    % \item Stable model that outperforms state-of-the-art contrastive learning models when batch size decreases.\todoff{add motivation in the intro about why robustness to a smaller batch size is important}

\end{enumerate}

% \todoffinline{Super pet peeve of mine: except for books or theses, don't take up space with this roadmap. Instead, in the above contributions, add pointers to the sections (or even subsections) where we provide evidence for each contribution}
% \todokdinline{I think Cynthia added this part for the last submission since we were short a few sentences to reach the page limit. We can remove this.}
% The remainder of this paper is organized as follows. In \cref{sec:Related-Work}, we describe existing work on visuolinguistic retrieval, contrastive learning, and multimodal learning. In \cref{sec:Problem-Description}, we formalize the object retrieval problem and give examples of the data under consideration. In \cref{sec:Method}, we describe our novel approach to learning models of linguistically-driven queries against a multimodal set of objects. Finally, in \cref{sec:Experiments}, we demonstrate the effectiveness of our approach against traditional contrastive learning and supervised contrastive learning baselines, both for the general learning problem and in the case of missing information at test time.

% \todocmi{NOTES: Some things are probably less informative or more ambiguous (e.g., depth for differentiating round things). So instead of predicting a single vector output, predict a vector and a covariance that captures uncertainty; could potentially use that to do more accurate search, or to handle really uncertain modalities (e.g., just drop speech when it's uninformative).}

%===================================================================

\section{Related Work}
\label{sec:Related-Work}

\todoffinline{use paragraphs to add some light structure to this block-o-text}

\todoffinline{I'm not convinced this section shouldn't appear toward the end of the paper. Maybe adding that narrative structure will help?}

There is extensive work on the task of image retrieval, of which physical object retrieval can be considered a special case. In this task, language is used to formulate queries against datasets of images, for example in text-and-image matching tasks for fashion data~\cite{gao2020fashionbert, wen2021comprehensive}, sketch retrieval~\cite{huang2017deep}, and general photographs of objects~\cite{ma2020large, novak2015large, hong2021gilbert}. Prior work has focused solely on language and visual modalities, using text or a combination of language and vision to perform visual retrieval~\cite{vo2019composing}. While recent works have focused on grounding with respect to a knowledge set \cite{10.1145/3357384.3357889,10.1145/3397271.3401097}, our work focuses on robust multimodal learning to perform grounding over an arbitrary number of input modalities.

The growing number of datasets that contain modalities beyond images and text demonstrates the importance and the challenge of this task. \citet{bagher-zadeh-etal-2018-multimodal} introduce a large dataset of multimodal opinion sentiment and emotion intensity (CMU-MOSEI) for sentiment analysis and emotion recognition. They also introduce a novel multimodal fusion technique called the Dynamic Fusion Graph (DFG).~\citet{GoLD_UMBC} present a multimodal dataset of household objects containing RGB images, depth images, written language, and spoken language, which has been used to support learning grounded language directly from speech given small datasets~\cite{KebeAAAI2022}. 
\citet{baltrusaitisMultimodalMachineLearning2019} propose a new taxonomy of multimodal machine learning by introducing five technical challenges in addition to typical early and late fusion categorization. They cover different approaches to multimodal machine learning such as \textit{representation}, \textit{translation}, \textit{alignment}, \textit{fusion}, and \textit{co-learning}. In this work, we focus on alignment, the process of identifying the direct relationship between elements across modalities.

We highlight a key difference with recent work like \citet{10.1145/3397271.3401232} and \citet{10.1145/3331184.3331213} who develop multi-modal retrieval models. These retrieval models are class based, where the model is trained to recognize any object of the same retrieval class as being ``same''. Our method and data is instance based, where different objects are not considered ``same'' even if they have the same class, due to the grounding goals of our work (i.e., the agent should identify the specified object, not equivalent objects). These works also do not consider the failure of a modality that is our target of interest.

A number of different approaches have been proposed for linguistic-based alignment learning. \citet{alayrac2020self} use a self-supervised contrastive learning method to learn and combine representations from three modalities of visual, audio, and language. Audio and visual domains are mapped to the same space because they are fine-grained, containing dense information for each frame of video. These embeddings are then mapped to a coarse space where text is also mapped. Their network respects specificity and comparability of different modalities. The loss function they define consists of two distance-based terms for the corresponding spaces, differing from ours in that they do not handle arbitrarily many modalities and do not focus on robustness in the face of modality drop-outs. \citet{Nguyen-RSS-20} take a similar approach and perform pairwise cosine similarity to align images and natural language text descriptions to perform object retrieval tasks. \citet{triplet_loss_2021_CVPR} take a cross-modal manifold alignment approach to grounded language learning using triplet loss \cite{Chechik:2010:LSO:1756006.1756042}, and perform object retrieval by connecting text descriptions of objects to their corresponding RGB-D images.

In general there are two different family of approaches to multimodal learning in the literature. One family is based on classification and cross-entropy losses such as~\citet{NEURIPS2020_supervised_contrastive, chen2020simple} and the other family is geometric based such as~\citet{ICML22GeometricMultimodal,Carvalho-cooking-triplet,salvador2017cooking,triplet_loss_2021_CVPR}.
In this paper we propose to marry the two families and show that combining geometric approach with cross-entropy method is superior. \todokdinline{Pun intended :-)}
\todokdinline{Should we move the following two paragraphs to the Approach section?}
For the classification part of our loss function we employ the supervised contrastive loss proposed by~\citet{NEURIPS2020_supervised_contrastive}. The way we formulate our geometric part of the loss function is similar to lifted structured loss~\cite{songCVPR16LiftedStructured}, where they take a geometric approach in a unimodal scenario and consider pairwise distances among all items in a batch, but in this paper we apply our idea to four modalities, and our formulation can be extended to any number of modalities.
This extended geometric formulation results in a loss function that is similar to \citet{ICML22GeometricMultimodal} where they propose a geometric contrastive learning approach in which they %use the same instance of an MLP network (same weights) after extracting embeddings from each modality, while we use different instances of the same MLP network (different weights). They also 
fuse all embeddings as a central embedding and then minimize the distance between each embedding and that central embedding. However, their method does not take into account any notion of classification.

The geometric approach is also known as contrastive learning~\cite{qin2021world}, which is based on the idea of similarity and distance of elements of training data~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR,salvador2017cooking}. Triplet loss is also a special case of contrastive loss~\cite{NEURIPS2020_supervised_contrastive}. Contrastive loss is mostly used in self-supervised learning~\cite{bui2021self, alayrac2020self,chen2020simple}, with some exceptions such as \citet{NEURIPS2020_supervised_contrastive}, who use contrastive loss for a supervised problem. Most of these methods are implemented for a non-multimodal dataset (e.g., RGB images only). Our work is a specialization of contrastive loss that is focused on the multimodal input and- language problem.

We note in particular that standard triplet learning approaches often require an expensive \textit{mining} step to find harder negative samples to enable effective learning \cite{Hoffer2015,Schroff2015,DBLP:conf/eccv/2018-9,DBLP:conf/eccv/ZhaoJQLH18,Zhai2018}. This is problematic from throughput and scalability, as run time becomes quadratic in batch size. Our approach does not require mining to obtain good performance, alleviating these issues along with their complicated history in reproducibility and parameter tuning \cite{Musgrave2020,Raff2020c,Raff2019_quantify_repro}.

Some of the most closely related work focuses on learning based on more than two input modalities. \citet{Veit_2018_CVPR} train a joint model of images, hashtags, and users to perform image retrieval. Their task is similar to ours---given a description, we want to find the image/object in the scene that best matches the description. They tackle this problem by forming a three-way tensor product model. They use a ranking loss to train the model where the score of an observed triplet is higher than an unobserved triplet. They sample six negative triplets per positive sample triplet, and use each of them as a negative in the loss. The downstream retrieval task is then simply done by taking the argmax of the tensor product for a given user. This work differs from ours in that it becomes computationally complex as more and higher-dimensional modalities are added.
Our proposed geometric loss function resembles work on quadruplet loss~\cite{chen2017beyond,tursun2021efficient} but is intended to scale to an arbitrary number of modalities.

\citet{semihet_three_way_Lei_2020} use inputs from three modalities of image, sketch, and edgemap to perform image retrieval given the sketch. They use co-attention mechanism, and their loss function contains alignment loss, cross-entropy loss, and sketch-edgemap contrastive loss. \citet{Mittal2020M3ER} use Canonical Correlational Analysis (CCA) to differentiate between ineffective and effective modalities for the emotion recognition task from multiple input modalities including face, speech, and text. Multimodal learning has been used for different applications including but not limited to object retrieval. \citet{Deception_ICMI_2014} use language, physiological response, and thermal sensing to detect deceit. \citet{het_data_fusion_liu_IEEE_2017} take two modalities as input and output predictions in a third modality. \citet{tursun2021efficient} uses quadruplet loss for two modalities only; image and sketch. A quadruplet is composed of a sketch picture as an anchor, a negative example from sketch domain, a negative example from picture domain,  and a positive example from picture domain. \citet{chen2017beyond} also uses quadruplet loss, but does not support multiple domains. In contrast, our method is not restricted to two modalities.
% \todo[inline]{talk about CLIP and DALL-E here}
To the best of our knowledge, there has been limited work on incorporating more than three modalities into a general learning approach.

%===================================================================

\section{Problem Description}
\label{sec:Problem-Description}

Given a language command, for example either text or speech, that describes an object, we want our model to retrieve the correct object from a set of objects. This problem is an exemplar of tasks found in the area of grounded language learning in the fields of robotics and natural  language processing. Intuitively, the goal is to take unconstrained natural language queries and select the appropriate object based on the complete set of sensor inputs available to the agent. We demonstrate on a domain containing four modalities, all of which refer to objects in the environment: spoken language, written text, RGB (image) inputs, and depth camera inputs. \Cref{fig:experimental-setup} illustrates a small visualization of our object retrieval task: the spoken query ``A white textbook titled algorithms'' is provided to our contrastive model, which identifies the item (outlined in red, and represented by up to four different modalities) as the most likely item the query is referring to.

\begin{figure}[tb]
\centering
\includegraphics[width=.5\columnwidth]{Figures/experiment-setup.png}
\caption{
The experimental object retrieval setup, in which objects are represented by four modalities: an RGB image, a depth image, spoken descriptions, and textual descriptions. Given a query in some modality, our approach seeks to select the object that is the best fit, per a trained model. This approach, detailed in \cref{sec:Method}, is able to rank objects as to appropriateness even when one or more modalities is ablated at test time (e.g., depth inputs are missing), and outperforms state-of-the-art contrastive learning approaches on this task.
}
\label{fig:experimental-setup}
\end{figure}

More formally, we consider a spoken language command $x_s$, a textual language command $x_t$, a set of RGB images $X_r = \{x_r^{(1..n)}\}$, and a set of depth images $X_d = \{x_d^{(1..n)}\}$, the task is to retrieve the correct object by choosing the index that has the minimum distance from either of the language commands across all modalities.
Depending on which modalities are or are not ablated, we consider up-to four distances: ``sr,'' a vector of distances between $x_s$ and all RGB images in $X_r$; ``sd,'' a vector of distances between $x_s$ and all depth images in $X_d$; ``tr,'' a vector of distances between $x_t$ and all RGB images in $X_r$; and ``td,'' a vector of distances between $x_t$ and all depth images in $X_d$. In order to select the correct object, we first perform a component-wise average of the relevant modality pair distances for the available modalities. Then, we select the object which had the minimum distance, i.e., we perform an argmin on this average vector of multiple-modality distances. Depending on the missing/available sensors during test time, we might have any combination of these four distances. For example, if no written instructions are available at test time,\footnote{This setting is particularly salient. While large bodies of text are frequently available at training time, a person interacting directly with a physical agent may well prefer to use spoken instructions.} we compute the component-wise average of $sr$ and $sd$, and then select the object whose coordinate resulted in the lowest average distance. This method allows us to extend our model to support arbitrary modalities while remaining robust to test cases in which some modalities are missing or incomplete.

%===================================================================

\section{Approach}
\label{sec:Method}

In keeping with previous work on the closely related problem of image retrieval, we focus on contrastive loss approaches, in which the goal is to learn an embedding of data where similar samples---in our case, samples belonging to the same class of object---are embedded `close' to one another, while dissimilar samples are farther apart. We develop a novel geometric loss function that simultaneously minimizes intra-class distances and maximizes inter-class distances across each pair of modalities yielding a model that is both effective at the retrieval task defined above and robust to modality dropouts at test time. We take an additional step and combine this geometric loss with a classification-based (cross-entropy) loss function which results in a superior model compared to either of geometric or cross-entropy losses alone.

\subsection{Geometric Alignment Loss}
\label{ssec:distanceloss}

% For $M$ modalities, we 
We define a distance-based loss function which can be used for arbitrary number of modalities. Our proposed method is inspired by the well-known similarity-based triplet loss~\cite{Carvalho-cooking-triplet,triplet_loss_2021_CVPR}, and is similar to contrastive loss~\cite{chen2020simple,NEURIPS2020_supervised_contrastive} under some settings.
Triplet loss-based learning works by forcing similar concepts from different domains `together' in some shared embedding space, while forcing dissimilar concepts `apart.' It is so named because it relies on three data points from the training set: a positive, a negative, and an anchor point. 
However, standard triplet loss cannot be used for more than two modalities. Therefore, we modify this concept as follows. During training, we sample two different instances and their corresponding representations from all modalities into two sets---one positive set (referring to a specific object) and one negative set (referring to some different object) as shown in \cref{fig:emma-loss}.
Unlike some prior triplet loss methods~\cite{GoLD_UMBC,triplet_loss_2021_CVPR}, the anchor is not randomly chosen from different modalities in each batch. Instead, in our setting, every item in the positive set becomes anchor once, and we minimize the distance between that item and other items in the positive set, while minimizing the distance between that item and all items in the negative set. It can be seen as an one-to-many relationship instead of an one-to-two relationship in the triplet loss formulation.
% the anchor and positive items are from the same object class, but in different modalities and each of the positives modalities becomes anchor once. 
We consider the following terminology:
% \todokdinline{I feel like using the terminology of triplet loss especially \textbf{anchor} is confusing here or maybe it needs to be rephrased. We can alternatively say that we consider all pairwise distances.}
% \todoffinline{I don't see the problem with anchor, but if you'd rather use a different term, what about query? As a note, while you do consider all pairwise distances but you treat some differently than others.}
\begin{itemize}
    % \item data point: a row in the csv/tsv file.
    \item Positive (Instance): A set of embeddings of one data point (e.g., an RGB image of an apple, corresponding depth image, text description, and speech description of the same apple)
    \item Negative (Instance): A set of embeddings of another data point of a different object (e.g., an RGB image of a book, corresponding depth image, text description, and speech description of the same book)
    \item Anchor (Modality): Every modality of the positive set is chosen as the learning anchor once. In our formulation, anchor is a ``concept'' referring to the points in the positive sets rather than a single instance in itself.
    %For our experiments, text is used. 
    The anchor is used as the basis for learning distances between positive and negative samples.
    % variance of text is lower and clusters of different objects are more distinct compared to other modalities, therefore, it can train a better model.
    % where we don't minimize the distance between the pair of embeddings from all modalities of the same data point, and we don't maximize the distance between positive and negative datapoints of the same modality.
    % \todo{explain this better}
    % \todokd{I rephrased it. Good?}
\end{itemize}

The objective is then to first minimize the distance between each pair of positive points from heterogeneous modalities, and second, maximize the distance between each positive and negative points from all modalities.
%and finally minimize the distance between each pair of negative points from heterogeneous modalities.
% The objective is then to first minimize the distance between the anchor and each of the other positive points from heterogeneous modalities, and second, maximize the distance between the anchor and each of the negative points from all modalities.
%and finally, maximize the distance between positive and negative anchor points from the same modality. 
% When we minimize the distance between each pair of positives, and maximize the distance between each pair of modalities one from the positive set and one from the negative set, the model converges faster than \supcon{} while maintaining the same performance when we have all modalities and a little bit less than Supcon when we drop text. 
We refer to this approach as \geom{}, for extended multimodal alignment which is formulated in \cref{eq:full-emma}. An illustration of this loss function is provided in \cref{fig:emma-loss}.

% \begin{equation}\label{eq:full-emma}
%     L = \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ \max (\cos(z_{m_1}^{+}, z_{m_2}^{-}),0) \right] + \sum_{m_3=m_1+1}^{M} \left[ 1 - \max(\cos(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right]
% \end{equation}
\begin{equation}\label{eq:full-emma}
    L = \sum_{m_1=1}^{M} \left[ \sum_{m_2=1}^{M} \left[ \max (\cos(z_{m_1}^{+}, z_{m_2}^{-}) -1 + \alpha,0) \right] + \sum_{m_3=m_1+1}^{M} \left[ \max(1 - \cos(z_{m_1}^{+},z_{m_3}^{+}),0) \right] \right]
\end{equation}

>>>>

\begin{table*}[tb]
    \centering
    \begin{subtable}[t]{1.00\textwidth}
    \centering
    \resizebox{.99\textwidth}{!}{
    \csvautotabular{resultsMRR.csv}
    }
    \end{subtable}
    \vskip\baselineskip
    % \hfill
    \begin{subtable}[t]{1.00\textwidth}
    \centering
    \resizebox{.99\textwidth}{!}{
    \csvautotabular{resultsACC.csv}
    }
    \end{subtable}
\end{table*}
% \todocm{For Table 1 - change all mrr.xx to MRR \textit{xx} for consistency with eqn 8, and all acc.xx to Acc \textit{xx}---e.g., column headers should be 'MRR \textit{td}'. Also make sure the caption edits are correct.}
% \todo[inline]{Make it two subtables, and remove MRR and Acc from the columns, and spell out sr, tsrd, and ...}

% Automatic DataTool Method
\DTLloaddb{resultsMRR}{resultsMRR.csv}
\DTLloaddb{resultsACC}{resultsACC.csv}
\begin{table}[]
    \centering
    \resizebox{1.0\textwidth}{!}{
    \DTLdisplaydb{resultsMRR}
    }
    \resizebox{1.0\textwidth}{!}{
    \DTLdisplaydb{resultsACC}
    }
    \caption{Caption}
    \label{table:quantitative}
\end{table}

% Manually constructed table
\begin{table*}[]
\centering
\resizebox{.97\textwidth}{!}{
\begin{tabular}{lc|ll|lll}
\multicolumn{1}{c}{} &  & \multicolumn{2}{c|}{\textbf{Accuracy}} & \multicolumn{3}{c}{\textbf{Mean Reciprocal Rank}} \\ \hline
\multicolumn{1}{c|}{\textbf{Method}} & \textbf{Modalities} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Speech/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/RGB/\\ Depth\end{tabular}}} & \multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Text/Speech/\\ RGB/Depth\end{tabular}}} \\ \hline
\multicolumn{1}{l|}{EMMA} & TSRD & \multicolumn{1}{l|}{\textbf{0.65}$\pm$0.01} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{\textbf{0.79}$\pm$0.002} & \multicolumn{1}{l|}{0.90$\pm$0.003} & 0.90$\pm$0.004 \\
\multicolumn{1}{l|}{EMMA} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.004} & -- \\
\multicolumn{1}{l|}{\supcon{}} & TSRD & \multicolumn{1}{l|}{0.59$\pm$0.007} & 0.84$\pm$0.008 & \multicolumn{1}{l|}{0.75$\pm$0.004} & \multicolumn{1}{l|}{0.91$\pm$0.006} & 0.90$\pm$0.006 \\
\multicolumn{1}{l|}{\supcon{}} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.006 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.004} & -- \\
\multicolumn{1}{l|}{Contrastive Cosine} & TSRD & \multicolumn{1}{l|}{0.40$\pm$0.153} & 0.83$\pm$0.006 & \multicolumn{1}{l|}{0.62$\pm$0.034} & \multicolumn{1}{l|}{0.90$\pm$0.004} & 0.89$\pm$0.005 \\
\multicolumn{1}{l|}{Contrastive Cosine} & 3 & \multicolumn{1}{l|}{--} & 0.84$\pm$0.005 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.91$\pm$0.003} & -- \\
\multicolumn{1}{l|}{\ours{}} & TSRD & \multicolumn{1}{l|}{--} & 0.83$\pm$0.008 & \multicolumn{1}{l|}{--} & \multicolumn{1}{l|}{0.90$\pm$0.005} & -- \\ \hline
\end{tabular}
}
\caption{Average and standard deviation of accuracy and MRR over 3 runs with 3 different random seeds on a held-out test set. These results represent performance when no modalities are ablated, in which case supervised contrastive loss, contrastive loss, and EMMA perform similarly (``--'' is an ablated modality and can't be evaluated). All metrics are from 0 to 1, and higher is better. For 5 objects, a random guess would have MRR of 0.33, and the worst case performance would be always ranking the correct item last which gives us 0.2. All models are trained using text as anchor. The batch size is 64 and optimizer is SGD for all experiments except the triplet loss, where batch size is 1 and optimizer is Adam.
%In the Triplet experiment, RGB and Depth embeddings are concatenated to form a single vision embedding since triplet loss cannot be used for more than two modalities.
} 
\label{table:quantitative}
\end{table*}
% \todo[inline]{This table is not referred to until sec:results, which is waaay down from where it's introduced, and it takes some parsing to figure out what's going on. Either add a paragraph under it that starts with something like "Table 1 shows high-level results, in which..." or just move the table to the results section.}

% %Edward moved table above where refernced b/c LaTeX plays games with double-wide objects
% \begin{table*}[bth]
% \centering
% \begin{tabular}{c|c|c|c|c|c|c}
% \toprule
% \textbf{Method} & \textbf{Modalities} & \textbf{acc.srd} & \textbf{acc.trd} & \textbf{MRR.srd} & \textbf{MRR.trd} & \textbf{MRR.tsrd} \\ %\hline
% \midrule
% EMMA (Text) & 4 & \textbf{0.65}$\pm$0.01 & 0.83$\pm$0.006 & \textbf{0.79}$\pm$0.002 & 0.90$\pm$0.003 & 0.90$\pm$0.004 \\
% EMMA (Text) & 3 & - & 0.84$\pm$0.008 & - & 0.90$\pm$0.004 & - \\
% % Simple MMA (Text)  & 3 & ?$\pm$? & ?$\pm$? & ? & ? \\
% % Simple MMA (Text) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
% % Simple MMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & ? & ?  \\
% DONE Supervised Contrastive (Text)& 4 & 0.59$\pm$0.007 & 0.84$\pm$0.008 & 0.75$\pm$0.004 & 0.91$\pm$0.006 & 0.90$\pm$0.006  \\
% TODO Supervised Contrastive (Text)& 3 & - & 0.84$\pm$0.006 & - & 0.91$\pm$0.004 & -\\
% DONE Contrastive Cosine (Text) & 4 & 0.40$\pm$0.153 & 0.83$\pm$0.006 & 0.62$\pm$0.034  & 0.90$\pm$0.004 & 0.89$\pm$0.005\\
% TODO Contrastive Cosine (Text) & 3 & - & 0.84$\pm$0.005 & -  & 0.91$\pm$0.003 & -\\
% PROJ Triplet (Text) & 3 & - & 0.83$\pm$0.008 & - & 0.90$\pm$0.005 & -  \\
% % DONE EMMA (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & - \\
% % DONE EMMA (Speech) & 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
% % DONE Supervised Contrastive (Speech)& 3 & ?$\pm$? & ?$\pm$? & - & -\\
% % DONE Supervised Contrastive (Speech)& 4 & ?$\pm$? & ?$\pm$? & ? & ? \\
% % TODO Contrastive Cosine (Speech) & 3 & ?$\pm$? & ?$\pm$?  & - & -\\
% % TODO Contrastive Cosine (Speech) & 4 & ?$\pm$? & ?$\pm$?  & ? & ?\\
% % DONE Triplet (Speech) & 3 & ?$\pm$? & ?$\pm$? & - & -  \\
